Augmentation of a Model Historical Harpsichord Mechanism for Haptic Enabled Interaction

### 2025-01-18 src/article.tex
footnote urls

The project look at restoring tactile and haptic response of 17th
century harpsichords for an exhibition at the musical instrument museum
San Colombano in Bologna. San Colombano hosts the Tagliavini Collection
(from Luigi Ferdinando Tagliavini), a collection of approximately 70
historical keyboard instruments in working order \cite{Tagliavini2007}
(https://genusbononiae.it/en/san-colombano-tagliavini-collection/). Some
items in the collection are not in playing condition or are too fragile
to continue to be available for the public to play. One of those fragile
instruments in the Tagliavini collection is a 1547 Harpsichord by
Alessandro Trasuntino \cite{Wraight2024}, which was used as the basis
for an interactive exhibition commisioned by San Colombano.

@@

https://github.com/Nemus-Project/haptic-harpsichord-nime-2025/issues/Nemus-Project/haptic-harpsichord-nime-2025/images/disegno.pdf

@@

\begin{itemize}
\item
  Nano form factor small
\item
  allows for easy changes in chipset
\item
  ``Native USB''
  https://learn.adafruit.com/dude-where-s-my-com-port/native-usb-boards\#
\item
  BLE Model

@@

To address this, once the wiring for the controller board was finalised
a new design was drafted this time using JST-PH connectors. No
commercial cables were immediately available in the configuration and
length required, as such a cable loom needed to be made which meant
crimping cables by hand. A crimping tool from JST is prohibitively
expensive compared to the cost of other components in the project*. The
crimping tool also appeared to be uncommon and was not readily available
in the workshops at the projects disposal. Eventually a tool was found,
but an alternative approach was also researched in order to maintain the
criteria of accessibility.

A locator was adapted from a creative commons stl model
https://www.thingiverse.com/thing:1646016 that allowed a generic
crimping tool to be adapted for use with JST-PH crimps†
\textbackslash figure\{\}. The locator was printed using a Bambu X1 3D
printer.

\emph{** A second alterntaive was assembling cables from other sets that
were at hand.} \emph{* from RS
https://uk.rs-online.com/web/p/crimp-tools/6880877 £508 inc VAT} \emph{†
model available in the git repository}

@@

The thresholds were calibrated visually utilising the serial plotter of
the Arduino IDE and MIDI was confirmed using the open source MIDI
Monitor (https://github.com/krevis/MIDIApps).

>>>>

The project look at restoring tactile and haptic response of 17th
century harpsichords for an exhibition at the musical instrument museum
San Colombano in Bologna. San Colombano hosts the Tagliavini Collection
(from Luigi Ferdinando Tagliavini), a collection of approximately 70
historical keyboard instruments in working order \cite{Tagliavini2007}
\footnote{https://genusbononiae.it/en/san-colombano-tagliavini-collection/}.
Some items in the collection are not in playing condition or are too
fragile to continue to be available for the public to play. One of those
fragile instruments in the Tagliavini collection is a 1547 Harpsichord
by Alessandro Trasuntino \cite{Wraight2024}, which was used as the basis
for an interactive exhibition commisioned by San Colombano.

@@

\footnote{https://github.com/Nemus-Project/haptic-harpsichord-nime-2025/issues/Nemus-Project/haptic-harpsichord-nime-2025/images/disegno.pdf}

@@

\begin{itemize}
\item
  Nano form factor small
\item
  allows for easy changes in chipset
\item
  ``Native USB''
  \footnote{https://learn.adafruit.com/dude-where-s-my-com-port/native-usb-boards}
\item
  BLE Model

@@

To address this, once the wiring for the controller board was finalised
a new design was drafted this time using JST-PH connectors. No
commercial cables were immediately available in the configuration and
length required, as such a cable loom needed to be made which meant
crimping cables by hand. A crimping tool from JST is prohibitively
expensive compared to the cost of other components in the project
\footnote{from RS https://uk.rs-online.com/web/p/crimp-tools/6880877
£508 (inc. VAT)}. The crimping tool also appeared to be uncommon and was
not readily available in the workshops at the projects disposal.
Eventually a tool was found, but an alternative approach was also
researched in order to maintain the criteria of accessibility.

A locator was adapted from a creative commons stl model
\footnote{https://www.thingiverse.com/thing:1646016} that allowed a
generic crimping tool to be adapted for use with JST-PH crimps†
\textbackslash figure\{\}. The locator was printed using a Bambu X1 3D
printer. A second alterntaive was assembling cables from other sets that
were at hand.

@@

The thresholds were calibrated visually utilising the serial plotter of
the Arduino IDE and MIDI was confirmed using the open source MIDI
Monitor \footnote{https://github.com/krevis/MIDIApps}.
